The LSST-DESC 3x2pt Tomography Optimization Challenge

### 2021-05-14 paper/paper.tex
address markus and shahab notes

The \emph{3x2pt} method has become a standard method for performing this combination.
In this method, two-point correlations are computed among and between two samples, the shapes of 
background (source) galaxies and the locations of foreground galaxies, which trace foreground
dark matter haloes.  The three combinations (source-source, source-lens, and lens-lens) are
measured in either Fourier or configuration space, and can be predicted from a combination of 
perturbation theory and simulation results.  The method has been used in the Dark Energy Survey, DES, 
\citep{des-3x2pt}, and to combine Baryon Oscillation Spectroscopic Survey, BOSS, with the Kilo-Degree 
Survey, KiDS \citep{kids-3x2pt}.

@@

Various recent work have explored general tomographic binning strategies. The general question of optimization
was recently discussed in detail
in \citet{rainbow} using a self-organizing map approach to target up to five tomographic bins.  For this
configuration they find that equally spaced redshift bins are a good proxy
for an optimized method, but note, importantly, that we should not be constrained to trying to directly
match our bin selections to specific tomographic bin edges. \citet{taylor18} considered using
fine tomographic bins as an alternative to fully 3D lensing, and found that the strategy of equal numbers
per bin is less effective at high redshift, a result we will echo in this paper. The fine-binning strategy
also allows a set of nulling methods to be applied, offering various advantages in avoiding poorly-understood
regimes \citep{taylor18b,bnt,xcut}.

This paper is part of preparations for the analysis of Legacy Survey of Space and Time (LSST) data at the Rubin Observatory, being run by the Dark Energy Science Collaboration (DESC).
In it, we discuss and evaluate the related challenge
of using a limited set of colour bands to make those assignments, motivated by requirements of using the
metacalibration method for galaxy shape measurement to limit biases.  
A similar question, with different motivations, was explored in \citet{jain}, who find that 3-band 
gri tomography could be effective for tomographic selection, a result we will echo here in a new context: in this paper we describe the results of a 
challenge using simulated Rubin-like data with a range of methods submitted by different teams.  We explore how many tomographic bins we can effectively generate using only a subset of bins, and compare methods submitted to the challenge.

In \autoref{sec:motivation} we explain the need for methods that work on limited sets of bands in the context of upcoming lensing surveys. In \autoref{sec:data} we describe the simulated data used in the challenge.
Section \ref{sec:results} briefly describes the entrants to the challenge, and discusses their performance (full
method details are given in appendix \ref{app:methods}).
 We conclude in \autoref{sec:conclusion}.

@@

Furthermore, metacal provides a solution to the pernicious and general problem of
\emph{selection biases} in WL.  Measurements of galaxy shear are known to have noise-induced errors
that are
highly covariant with those on other galaxy properties, including size and, most importantly, flux and
magnitude.  It follows that a cut (or re-weighting) on galaxy magnitudes, or on any quantity derived
from  them, will induce a shear bias since galaxies will preferentially fall either side of the cut
depending on their shear.  Within metacal, we can compute and thus account for this bias
very accurately, by performing the cut on both
sheared variants of the catalogue, and measuring the difference between the post-cut shear in the two
variants.  In DES the corrected biases were found up to around 4\%, far larger than
the requirements for the current generation of surveys \citep{des-y1-cat}.

In summary: metacal allows us to correct for significant selection biases, but only if all selections
are performed using only bands in which the PSF is measured well enough for deconvolution to be
performed.  For Rubin, this means using only the $r$, $i$, $z$, and perhaps $g$ bands.  In this work we
therefore study how well we can perform tomography using only this limited photometric information.  In
particular, this limitation prevent us from using the most obvious method for tomography,
and computing a complete photometric probability density function (PDF) for each galaxy and assigning
a bin using the mean or peak of that PDF\footnote{This only prevents us using other photometry for \emph{selection}, not for \emph{characterization} after objects have been selected, such as computing the overall
number density $n(z)$ for a tomographic bin.}.

Simulation methods can also be used to correct for selection biases, provided that they match the
real data to high accuracy.  If we determine that limiting the bands we can use for tomography results
in a significant decrease in our overall signal-to-noise, outweighing the gain from the improved shape
calibration then this might suggest moving to rely more on such methods\footnote{One can account for
residual shear estimation uncertainty during parameter estimation by marginalizing over an unknown
factor.  Widening the prior on this factor decreases the overall constraining power of the analysis.}.

\section{Challenge Design}

The DESC Tomographic Challenge was launched in May 2020 and 
advertised among the LSST cosmology community, though it was open to outside entrants.

In the challenge, participants assigned simulated galaxies to tomographic bins,
using only their (g)riz bands and associated errors, and the galay size.  Their goal was to maximize the cosmological
information in the final split data set; this is generally acheived by separating different bins by 
redshift as much as possible.  Particpants were free to optimize or simply select nominal (target) 
redshift bin edges from their training sample, but in either case had to assign galaxies into the 
different bins.

@@

This differs from many redshift estimation questions because we are not at all concerned with estimates
of individual galaxy redshifts; instead, this challenge was desgined as a classification and bin-optimization
problem, amenable to more direct machine-learning methods.  As such we used the training / validation / testing
approach, as standard in machine-learning analyses.
The determination of the $n(z)$ of the recovered
bins (whether with per-galaxy estimates or otherwse) was not part of the challenge.

@@

The CosmoDC2 simulation was designed to support the DESC and 
is described in detail in \citet{cosmodc2}.  It covers 440
deg${}^2$ of sky, and used the dark matter particles from the Outer Rim
simulations \citep{outer_rim}.  The UniverseMachine \citep{universe_machine}
simulations were then used, with the GalSampler technique \citep{galsampler},
in a combination of empirical and semi-analytic
methods, to assign galaxies with a limited set of specified properties to halos.

@@

The Buzzard catalogue was developed to support DES Year 1 (DES-Y1)
data analysis, and is described in detail in \citet{buzzard}.
It has previously been used in DESC analyses, for example in \citet{dc1_pz}.
The catalogue used dark
matter simulations from L-GADGET2 \cite{gadget2} and then added galaxies using
an algorithm that matches a set of galaxy
property probability densities to data using sub-halo abundance matching. 
Buzzard galaxies extend to around redshift 2.3 (significantly shallower than CosmoDC2)
and were shown to be a good
match (after appropriate selections) to DES-Y1 observable catalogues.  Magnitudes
using the LSST band passes were provided in the catalogue.

@@

After this section, and cuts to contiguous geographic regions, we used around 20M objects from the Buzzard
simulations and around 36M objects the CosmoDC2 simulations.

@@

In this challenge we use only a lensing source (background) population, since the metacal requirements
do not apply to the foreground lens sample. We do, however, calculate 3x2pt metrics using the
same source and lens populations, both because this scenario is one that will be used in some analyses,
and because clustering-like statistics of lensing samples are important for studies of intrinsic
alignments. We use three different sets of metrics, one based on the over signal to noise of the spectra 
derived from the samples, and one based on a figure-of-merit for the constraints that would be obtained 
using them. The relationships between these metrics in our results is discussed in \autoref{sec:metric-results}.
We compute each metric on three different sets of summary statistics: lensing-alone, clustering-alone,
and the full 3x2pt. We therefore compute a total of nine values for each bin selection.

@@

Our first metric is the total signal-to-noise ratio (SNR) of the power spectrum derived from the assigned
$n_i(z)$ values:
\begin{equation}
    S = \mu^{T} C^{-1} \mu
\label{eq:snr}
\end{equation}
where $\mu$ is a stack of the theoretical predictions for the $C_\ell$ spectra for each tomographic 
bin, in turn, and $C$ a Gaussian estimate of the covariance between them, as described in
\citet{takada_jain} and summarized in \autoref{app:theory}.   We compute this metric for lensing alone, clustering alone, and the full 3x2pt combination.

@@

Metric 2 uses a figure of merit (FOM) based on that presented by the Dark Energy Task Force (DETF) \citep{detf}.  We 
use the inverse area of a Fisher Matrix ellipse, which approximates the size in two axes 
of the posterior PDF one would obtain on analysing the spectra:
\begin{align}
    F &= \left( \frac{\partial \mu}{\partial \theta} \right)^T C^{-1} \left( \frac{\partial \mu}{\partial \theta} \right), \\
    S &= \frac{1}{2 \pi \sqrt{\det{([F^{-1}]_{p_1, p_2})}}}
\label{eq:fom}
\end{align}
where $[F^{-1}]_{p_1, p_2}$ extracts the 2x2 matrix for two parameters.  We use both the original
DETF variant in which $p_1 = w_0$ and $p_2 = w_a$, representing the constraining power on dark energy 
evolution, and a version in which $p_1 = \Omega_c$ and $p_2 = \sigma_8$, representing constraining
power on overall cosmic structure amplitude and its evolution.

@@

The methods described above were run on a data set, previously unseen by entrants but drawn
from the same underlying population, of size 8.6M objects for CosmoDC2 and 5.4M objects for Buzzard.
A complete suite of the metrics described in Section \ref{sec:metrics} was run on each of the two
for each method.  In total, each method was run on 16 scenarios, each combination of: $griz$ vs. $riz$,
Buzzard vs. CosmoDC2, and 3, 5, 7, and 9 bins requested.  Nine metrics were calculated for each
scenario: clustering, lensing, and 3x2pt for each of the SNR metric (Equation \ref{eq:snr}), the $w_0-w_a$ 
DETF FOM, and the $\Omega_c - \sigma_8$ FOM (Equation \ref{eq:fom}).

@@

The overall winning method in the 3x2pt case in the fiducial $griz$ 3x2pt CosmoDC2 metric was \textsc{funbins},
which used a random forest to assign bins, but chose nominal bin edges by splitting the range spanned
by the sample into equal co-moving distance bins. As described in \autoref{sec:funbins},  this was designed
to minimze the shot noise for the angular power in each bin, and this has proven successful. 
Figure \ref{fig:funbin_nz} shows the number density obtained by this method, for both the riz and griz 
scenarios, and an additional post-challenge run using the full Rubin $ugrizy$ bands.  The metrics
obtained with \textsc{funbins} on $ugrizy$ are barely higher than those with $griz$ -- the largest
increase is 8\% for the $\Omega_c-\sigma_8$ metric, with the remainder all 5\% or less.

@@

For comparison, we also show results using all six LSST bands, $ugrizy$; these could not be used
with the metacalibration method. The improvement in the sharpness of the bins is noticeable but not large.
We quantify this in Figure \ref{fig:overlap}, which shows the fraction of objects that are in ambiguous
redshift slices, defined as objects in redshift ranges where there are more objects assigned to another 
tomographic bin.  The total fraction of such objects is also shown. This illustrates how adding the $g$ band 
improves the assigment
for almost every pair of bins, and reduce the ambiguous fraction by a factor of three to the 
point where a significant majority of galaxies are clearly assigned.  Further adding the remaining $u$ and $y$
bands improves the assignment in a way that is smaller in both relative and absolute terms.
This latter addition also does not translate into large increases in the figures of merit: the 
DETF score for CosmoDC2 is 169.7, barely higher than the score for $griz$, and this is also true for other 
metrics.   While this is only a general indication and not a thorough exploration, it is encouraging.

\begin{figure*}[htbp]
% \centering
\includegraphics[width=0.9\linewidth]{results/nzoverlap_dc2_funbins.pdf}
\caption{The pairwise fractions of total objects in redshift regions where multiple tomographic bins contain galaxies, for 
the \textsc{funbins} tomography method and three choices of bands.  
The $x$- and $y$-axes 
show tomographic bin indices, 
and the color represents the fraction of objects (compared to the total number in the challenge) that are 
at redshifts where both tomographic bins have members, in bins of width $\delta z=10^{-3}$.  
This metric corresponds to
the fraction of the area that is shaded in two colours in Figure \ref{fig:funbin_nz}.
The total-overlap figure gives the total fraction of objects
that are ambiguous over all bins, and is the sum of all the pairwise values.}
\label{fig:overlap}
\end{figure*}

@@

\subsection{Non-assignment}
The challenge did not require entrants to assign all galaxies to bins; aside from the impact of reduced final sample size on the metrics, no penalty was applied for discarding galaxies entirely.  Several methods
took advantage of this.  The ComplexSom method (see \autoref{sec:csom}) found
it did not improve scores, but as noted in \autoref{sec:pqnld} this may be an artifact
of the high density of the training data, and so could be explored in future challenges.

@@

This is not true of the overall winning method for large bin counts - as shown in the tables
in \autoref{app:tables}, significantly more methods acheive a top-scoring spot in the Buzzard
table \ref{tab:full_buzz} than in the CosmoDC2 table \ref{tab:full_dc2}. This reinforces the
conclusions made below in \autoref{sec:train-vs-fix}, though of course the relatively noisy
results make this difficult to interpret.

@@

\begin{figure}[htbp]
% \centering
\includegraphics[width=0.9\linewidth]{results/metric_comparisons.pdf}
\caption{A comparison of the metrics used in our challenge. The x-axis is shared, and shows
the 3x2pt $w_0 - w_a$ (DETF) figure of merit on the CosmoDC2 data set for all methods, 
with colours labelling the number
of bins used.  Each of the panels shows a metric which is the same except for a single change, 
which is labelled on its y-axis.  For example, the top panel still shows the 3x2pt $w_0 - w_a$ (DETF) figure of merit,
but for the Buzzard data set.}
\label{fig:metrics}
\end{figure}

@@

\begin{figure*}
\includegraphics[width=0.9\linewidth]{results/g_band_loss.pdf}
\caption{The degradation in constraining power over all methods
 when comparing them on methods including and excluding
the g-band, in terms of the ratio of the $w_0-w_a$ figure of merit.  The left-hand panel shows
results for the CosmoDC2 scenarios, and the right-hand panel for the Buzzard scenarios.  Colours
indicate the number of bins used in the analysis.}
\label{fig:loss}
\end{figure*}

@@

Some of the methods in the challenge accepted target bin edges as an input to their algorithms,
and then used classifiers to assign objects to those bins.  Other methods optimized the values
of bin edges themselves, by optimizing the metric.  For the runs shown here, these methods mostly
targeted one of the 3x2pt metrics.

@@

Among methods with fixed bins, those using equal numbers in each tend to plateau at scores of
120-140 in Table \ref{tab:cosmodc2}, suggesting that a range of machine learning models can
effectively classify by bin, at least in this case of representative, extensive training
information.  Notably, this is the case even for the \textsc{UTOPIA} entry, which uses the nearest 
training set neighbour to assign bins and thus is explicitly optimised
for this idealised scenario; it should give an upper limit when using the same fixed bins.
That other methods with the same nominal bin choice acheive scores close to it once again
confirms that many entries are close to the best possible score for this approach.

Several of the methods with optimized target bin edges break this
barrier and acheive very high scores.  This included ZotNet, ZotBin,  and JaxResNet.  The
Stacked Generalization method similarly included manually tuned bin edges to maximize
score for this scenario (and indeed its ensemble combination method could be applied to
combine the other successful methods). The typical 15--20\% improvement in FoM scores when optimizing
nominal bin edges makes clear that this easy win should be compelling.

@@

\section{Conclusions} \label{sec:conclusion}
The Tomography Challenge has proved a useful mechanism for encouraging the development
of methods for this science question.  We encourage this challenge structure for other
well-specified science cases, and DESC has subsequently run other challenges along the same lines
such is its N5K beyond-limber challege\footnote{\url{https://github.com/LSSTDESC/N5K}}.
We describe in Appendix \ref{sec:mistakes} some lessons we have learned from this challenge
that may be useful for future ones.

The results of the challenge have shown first that three or four-band photometry
is sufficient for specifying tomography even up to a large number of bins, if
the training data is complete enough: degneracies between colour do not make this impossible.
Excluding the g-band from the data reduces the constraining power by a noticeable but
not catastrophic amount.

@@

The challenge was heavily simplified, with the intention of testing whether
effective bin-assignments with limited photometry is possible even in theory,
and of generating as many potential methods for the task as possible.  With
both these acheived, future directions for simulating the problem are clear.
The most obvious is limited spectroscopic completeness and size.  Realistic
training data sets will be far smaller than the millions of objects we used here,
and the difficulty of measuring spectra for faint galaxies will make the them
incomplete and highly unrepresentative. Machine learning methods will be particularly
affected by the latter problem.  Future challenges must explore these important limitations
using simulated incompleteness.

@@

@@

We combine the Convolutional layers are with dimensionality reduction (pooling layers) and different
activation functions. Here, we used Autokeras to optimize a general convolutional architecture,
using the galaxies' magnitudes and colours to assign them to redshift bins, which were chosen such
that an equal number of galaxies were assigned to each.

@@

In addition to this base functionality, the SimpleSOM method includes a
mechanism for distilling $N_{\rm cell}$ SOM cells into $N_{\rm group}<N_{\rm
cell}$ groups of cells, while maintaining optimal redshift resolution. This is
done by invoking a hierarchical clustering of SOM cells \citep[see Appendix B of
][]{Wright/etal:2020a}, where in this case we combine cells based on the
similarity of the cells' $n(z) $moments (mean, median, and normalised median absolute
deviation from median) using complete-linkage clustering. The result of this
procedure is the construction of fewer discrete groups of training and reference
data, while not introducing pathological redshift-distribution broadening (as
happens when, e.g., training using a lower resolution SOM and/or clustering
cells based on distances in colour-magnitude space).

@@

We chose the learning algorithm for both methods to minimize the reciprocal of
the figure of merit $F$, where $F$ is defined in eqref{E:FOM}. We implement
these models using the JAX-based Flax neural network library \cite{jax}.

>>>>

The \emph{3x2pt} method has become a standard method for performing this combination.
In this method, two-point correlations are computed among and between two samples, the shapes of 
background (source) galaxies and the locations of foreground galaxies, which trace foreground
dark matter haloes.  The three combinations (source-source, source-lens, and lens-lens) are
measured in either Fourier or configuration space, and can be predicted from a combination of 
perturbation theory and simulation results.  The method has been used in the Dark Energy Survey, DES, 
\citep{des-3x2pt}, and to combine the Kilo-Degree 
Survey, KiDS with spectroscopic surveys \citep{kids-3x2pt, kids_gama, kids_2df}.

@@

Various recent work have explored general tomographic binning strategies. The general question of optimization
was recently discussed in detail
in \citet{rainbow} using a self-organizing map (SOM) approach to target up to five tomographic bins.  For this
configuration they find that equally spaced redshift bins are a good proxy
for an optimized method, but note, importantly, that we should not be constrained to trying to directly
match our bin selections to specific tomographic bin edges. 
They also highlight the utility of rejecting outlier or hard-to-classify objects from samples, and the ability of SOMs
to do so cleanly.
\citet{taylor18} considered using
fine tomographic bins as an alternative to fully 3D lensing, and found that the strategy of equal numbers
per bin is less effective at high redshift, a result we will echo in this paper. The fine-binning strategy
also allows a set of nulling methods to be applied, offering various advantages in avoiding poorly-understood
regimes \citep{taylor18b,bnt,xcut}.

This paper is part of preparations for the analysis of Legacy Survey of Space and Time (LSST) data at the Rubin Observatory, being run by the Dark Energy Science Collaboration (DESC).
In it, we discuss and evaluate the related challenge
of using a limited set of colour bands to make those assignments, motivated by requirements of using the
metacalibration method for galaxy shape measurement to limit biases.  
A similar question, with different motivations, was explored in \citet{jain}, who found that 3-band 
$gri$ tomography could be effective for tomographic selection, a result we will echo here in a new context: in this paper we describe the results of a 
challenge using simulated Rubin-like data with a range of methods submitted by different teams.  We explore how many tomographic bins we can effectively generate using only a subset of bins, and compare methods submitted to the challenge.

In \autoref{sec:motivation} we explain the need for methods that work on limited sets of bands in the context of upcoming lensing surveys. In \autoref{sec:design} we describe the design of the challenge and the simulated data used in it.
Section \ref{sec:results} briefly describes the entrants to the challenge, and discusses their performance (full
method details are given in appendix \ref{app:methods}).
 We conclude in \autoref{sec:conclusion}.

@@

Furthermore, metacal provides a solution to the pernicious and general problem of
\emph{selection biases} in WL.  Measurements of galaxy shear are known to have noise-induced errors
that are
highly covariant with those on other galaxy properties, including size and, most importantly, flux and
magnitude.  It follows that a cut (or re-weighting) on galaxy magnitudes, or on any quantity derived
from  them, will induce a shear bias since galaxies will preferentially fall on either side of the cut
depending on their shear.  Within metacal, we can compute and thus account for this bias
very accurately, by performing the cut on both
sheared variants of the catalogue, and measuring the difference between the post-cut shear in the two
variants.  In DES the corrected biases were found up to around 4\%, far larger than
the requirements for the current generation of surveys \citep{des-y1-cat}.

In summary: metacal allows us to correct for significant selection biases, but only if all selections
are performed using only bands in which the PSF is measured well enough for deconvolution to be
performed.  For Rubin, this means using only the $r$, $i$, $z$, and perhaps $g$ bands.  In this work we
therefore study how well we can perform tomography using only this limited photometric information.  In
particular, this limitation prevents us from using the most obvious method for tomography,
and computing a complete photometric probability density function (PDF) for each galaxy and assigning
a bin using the mean or peak of that PDF\footnote{This only prevents us using other photometry for \emph{selection}, not for \emph{characterization} after objects have been selected, such as computing the overall
number density $n(z)$ for a tomographic bin.}.

Simulation methods can also be used to correct for selection biases, provided that they match the
real data to high accuracy.  If we determine that limiting the bands we can use for tomography results
in a significant decrease in our overall 3x2pt signal-to-noise, outweighing the gain from the improved shape
calibration then this might suggest moving to rely more on such methods\footnote{One can account for
residual shear estimation uncertainty during parameter estimation by marginalizing over an unknown
factor.  Widening the prior on this factor decreases the overall constraining power of the analysis.}.
Constructing simulations with the required fidelity is challenging at Rubin-level accuracy and requires
careful comparison to deep field data.

\section{Challenge Design} \label{sec:design}

The DESC Tomographic Challenge opened in May 2020 and accepted entries until September 2020.

In the challenge, participants assigned simulated galaxies to tomographic bins,
using only their (g)riz bands and associated errors, and the galay size.  Their goal was to maximize the cosmological
information in the final split data set; this is generally achieved by separating different bins by 
redshift as much as possible.  Particpants were free to optimize or simply select nominal (target) 
redshift bin edges from their training sample, but in either case had to assign galaxies into the 
different bins.

@@

This differs from many redshift estimation questions because we are not at all concerned with estimates
of individual galaxy redshifts; instead, this challenge was desgined as a classification and bin-optimization
problem, amenable to more direct machine-learning methods.  As such we used the training / validation / testing
approach, as standard in machine-learning analyses.
The determination of the $n(z)$ of the recovered
bins (whether with per-galaxy estimates or otherwise) was not part of the challenge.

@@

The CosmoDC2 simulation was designed to support DESC and 
is described in detail in \citet{cosmodc2}.  It covers 440
deg${}^2$ of sky, and used the dark matter particles from the Outer Rim
simulations \citep{outer_rim}.  The UniverseMachine \citep{universe_machine}
simulations were then used, with the GalSampler technique \citep{galsampler},
in a combination of empirical and semi-analytic
methods, to assign galaxies with a limited set of specified properties to halos.

@@

The Buzzard catalogue was developed to support the DES Year 1 (DES-Y1)
data analysis, and is described in detail in \citet{buzzard}.
It has previously been used in DESC analyses, for example in \citet{dc1_pz}.
The catalogue used dark
matter simulations from L-GADGET2 \cite{gadget2} and then added galaxies using
an algorithm that matches a set of galaxy
property probability densities to data using sub-halo abundance matching. 
Buzzard galaxies extend to around redshift 2.3 (significantly shallower than CosmoDC2)
and were shown to be a good
match (after appropriate selections) to DES-Y1 observable catalogues.  Magnitudes
using the LSST band passes were provided in the catalogue.

@@

After this section, and cuts to contiguous geographic regions, we used around 20M objects from the Buzzard
simulations and around 36M objects from the CosmoDC2 simulations.

@@

In this challenge we use only a lensing source (background) population, since the metacal requirements
do not apply to the foreground lens sample. We do, however, calculate 3x2pt metrics using the
same source and lens populations, both because this scenario is one that will be used in some analyses,
and because clustering-like statistics of lensing samples are important for studies of intrinsic
alignments. We use three different sets of metrics, one based on the overall signal to noise of the spectra 
derived from the samples, and two based on a figure-of-merit for the constraints that would be obtained 
using them. The relationships between these metrics in our results is discussed in \autoref{sec:metric-results}.
We compute each metric on three different sets of summary statistics: lensing-alone, clustering-alone,
and the full 3x2pt. We therefore compute a total of nine values for each bin selection.

@@

Our first metric is the total signal-to-noise ratio (SNR) of the power spectrum derived from the assigned
$n_i(z)$ values:
\begin{equation}
    S = \mu^{T} C^{-1} \mu
\label{eq:snr}
\end{equation}
where $\mu$ is a stack of the theoretical predictions for the $C_\ell$ spectra for each tomographic 
bin, in turn, and $C$ a estimate of the covariance between them, making the approximation that the underlying density field has a Gaussian distribution as described in
\citet{takada_jain} and summarized in \autoref{app:theory}.   We compute this metric for lensing alone, clustering alone, and the full 3x2pt combination.

@@

Metric 2 uses a figure of merit (FOM) based on that presented by the Dark Energy Task Force (DETF) \citep{detf}.  We 
use the inverse area of a Fisher Matrix ellipse, which approximates the size in two axes 
of the posterior PDF one would obtain on analysing the spectra:
\begin{align}
    F &= \left( \frac{\partial \mu}{\partial \theta} \right)^T C^{-1} \left( \frac{\partial \mu}{\partial \theta} \right), \\
    S &= \frac{1}{2 \pi \sqrt{\det{([F^{-1}]_{p_1, p_2})}}}
\label{eq:fom}
\end{align}
where $[F^{-1}]_{p_1, p_2}$ extracts the 2x2 matrix for two parameters.  We use both the original
DETF variant in which $p_1 = w_0$ and $p_2 = w_a$, representing the constraining power on dark energy 
evolution, and a version in which $p_1 = \Omega_c$ and $p_2 = \sigma_8$, representing constraining
power on overall cosmic structure amplitude and its evolution. In each case the complete parameter space consists
of the eight $w_0w_aCDM$ cosmological parameters. This simplified measure does not include any nuisance
parameters modelling systematic errors in the analysis (most importantly the galaxy bias parameters in the clustering 
spectra), and like all Fisher analyses approximates posterior constraints as Gaussian.

@@

The methods described above were run on a data set, previously unseen by entrants but drawn
from the same underlying population, of size 8.6M objects for CosmoDC2 (our fiducial simulation) and 5.4M objects for Buzzard.
A complete suite of the metrics described in Section \ref{sec:metrics} was run on each of the two
for each method.  In total, each method was run on 16 scenarios, each combination of: $griz$ vs. $riz$,
Buzzard vs. CosmoDC2, and 3, 5, 7, and 9 bins requested.  Nine metrics were calculated for each
scenario: clustering, lensing, and 3x2pt for each of the SNR metric (Equation \ref{eq:snr}), the $w_0-w_a$ 
DETF FOM, and the $\Omega_c - \sigma_8$ FOM (Equation \ref{eq:fom}).

@@

Before the challenge began we somewhat arbitrarily chose 
the 9-bin 3x2pt $griz$ CosmoDC2 $w_0-w_a$ metric as the fiducial scenario on which
an overall winning method would be selected.
By this criterion the best method was, by a small margin, \textsc{funbins},
which used a random forest to assign bins, but chose nominal bin edges by splitting the range spanned
by the sample into equal co-moving distance bins. 
Other methods performed better in other branches of the challenge; in particular we note \textsc{ZotNet}, \textsc{Stacked Generalization}, and \textsc{NeuralNetwork} (see sections \ref{sec:zot}, \ref{sec:stackgen}, and \ref{sec:nn} respectively), which each won more than three 9-bin scenarios.

As described in \autoref{sec:funbins},  \textsc{funbins} was designed
to minimze the shot noise for the angular power in each bin, and this has proven successful. 
Figure \ref{fig:funbin_nz} shows the number density obtained by this method, for both the riz and griz 
scenarios, and an additional post-challenge run using the full Rubin $ugrizy$ bands.  The metrics
obtained with \textsc{funbins} on $ugrizy$ are barely higher than those with $griz$ -- the largest
increase is 8\% for the $\Omega_c-\sigma_8$ metric, with the remainder all 5\% or less.

@@

For comparison, we also show results using all six LSST bands, $ugrizy$; these could not be used
with the metacalibration method. The improvement in the sharpness of the bins is noticeable but not large.
We quantify this in Figure \ref{fig:overlap}, which shows the fraction of objects that are in ambiguous
redshift slices, defined as objects in redshift ranges where there are more objects assigned to another 
tomographic bin.  The total fraction of such objects is also shown. This illustrates how adding the $g$ band 
improves the assigment
for almost every pair of bins, and reduce the ambiguous fraction by a factor of three to the 
point where a significant majority of galaxies are clearly assigned.  Further adding the remaining $u$ and $y$
bands improves the assignment in a way that is smaller, though not insignificant, in both relative and absolute terms.
This latter addition also does not translate into large increases in the figures of merit: the 
DETF score for CosmoDC2 is 169.7, barely higher than its score of 167.2 for $griz$, and this is also true for other 
metrics.   While this is only a general indication and not a thorough exploration, it is encouraging.

\begin{figure*}[htbp]
% \centering
\includegraphics[width=0.9\linewidth]{results/nzoverlap_dc2_funbins.pdf}
\caption{The pairwise fractions $f_\mathrm{mis}$ of total objects in redshift regions where multiple tomographic bins contain galaxies, for 
the \textsc{funbins} tomography method and three choices of bands.  
The $x$- and $y$-axes 
show tomographic bin indices, 
and the color represents the fraction of objects (compared to the total number in the challenge) that are 
at redshifts where both tomographic bins have members, in bins of width $\delta z=10^{-3}$.  
This metric corresponds to
the fraction of the area that is shaded in two colours in Figure \ref{fig:funbin_nz}.
The total-overlap figure gives the total fraction of objects
that are ambiguous over all bins, and is the sum of all the pairwise values.}
\label{fig:overlap}
\end{figure*}

@@

\subsection{Non-assignment}
The challenge did not require entrants to assign all galaxies to bins; aside from the impact of reduced final sample size on the metrics, no additional penalty was applied for discarding galaxies entirely.  Several methods
took advantage of this.  The ComplexSom method (see \autoref{sec:csom}) found
it did not improve scores, but as noted in \autoref{sec:pqnld} this may be an artifact
of the high density of the training data, and so could be explored in future challenges\footnote{In real data or more realistic simulations discarding galaxies has a more complicated effects than simply decreasing sample size as it does here - removing objects with broad $p(z)$ can narrow the final $n(z)$, not just lower it. This was not a factor in this challenge because we use true redshifts to calculate our metrics.}.

@@

This is not true of the overall winning method for large bin counts - as shown in the tables
in \autoref{app:tables}, significantly more methods achieve a top-scoring spot in the Buzzard
table \ref{tab:full_buzz} than in the CosmoDC2 table \ref{tab:full_dc2}. This reinforces the
conclusions made below in \autoref{sec:train-vs-fix}, though of course the relatively noisy
results make this difficult to interpret.

@@

\begin{figure}[htbp]
% \centering
\includegraphics[width=0.9\linewidth]{results/metric_comparisons.pdf}
\caption{A comparison of the metrics used in our challenge for the different methods that were entered. The x-axis is shared, and shows
the 3x2pt $w_0 - w_a$ (DETF) figure of merit on the CosmoDC2 data set for all methods, 
with colours labelling the number
of bins used.  Each of the panels shows a metric which is the same except for a single change, 
which is labelled on its y-axis.  For example, the top panel still shows the 3x2pt $w_0 - w_a$ (DETF) figure of merit,
but for the Buzzard data set.}
\label{fig:metrics}
\end{figure}

@@

\begin{figure*}
\includegraphics[width=0.9\linewidth]{results/g_band_loss.pdf}
\caption{The degradation in constraining power over all methods
 when comparing them on methods including and excluding
the g-band, in terms of the ratio of the $w_0-w_a$ figure of merit.  The left-hand panel shows
results for the CosmoDC2 scenarios, and the right-hand panel for the Buzzard scenarios.  Colours
indicate the number of bins used in the analysis. Some methods perform worse when adding the $g$-band 
and have ratios greater than unity.}
\label{fig:loss}
\end{figure*}

@@

Some of the methods in the challenge accepted target bin edges as an input to their algorithms,
and then used classifiers to assign objects to those bins.  Entrants could select these edges; many followed the example random forest entry and split into equal counts per bin. Other methods optimized the values
of bin edges themselves, by maximizing a metric.  For the runs shown here, these methods mostly
targeted one of the 3x2pt metrics.

@@

Among methods with fixed bins, those using equal numbers in each tend to plateau at scores of
120-140 in Table \ref{tab:cosmodc2}, suggesting that a range of machine learning models can
effectively classify by bin, at least in this case of representative, extensive training
information.  Notably, this is the case even for the \textsc{UTOPIA} entry, which uses the nearest 
training set neighbour to assign bins and thus is explicitly optimised
for this idealised scenario; it should give an upper limit when using the same fixed bins.
That other methods with the same nominal bin choice achieve scores close to it once again
confirms that many entries are close to the best possible score for this approach.

Several of the methods with optimized target bin edges break this
barrier and achieve very high scores.  This included ZotNet, ZotBin,  and JaxResNet.  The
Stacked Generalization method similarly included manually tuned bin edges to maximize
score for this scenario (and indeed its ensemble combination method could be applied to
combine the other successful methods). The typical 15--20\% improvement in FoM scores when optimizing
nominal bin edges makes clear that this easy win should be compelling.

@@

\section{Conclusions} \label{sec:conclusion}
The Tomography Challenge has proved a useful mechanism for encouraging the development
of methods for this science question.  We encourage this challenge structure for other
well-specified science cases, and DESC has subsequently run other challenges along the same lines
such as the N5K beyond-limber challenge\footnote{\url{https://github.com/LSSTDESC/N5K}}.
We describe in Appendix \ref{sec:mistakes} some lessons we have learned from this challenge
that may be useful for future ones.

The results of the challenge have shown first that three or four-band photometry
is sufficient for specifying tomography even up to a large number of bins, if
the training data is complete enough: degeneracies between colour do not make this impossible.
Excluding the g-band from the data reduces the constraining power by a noticeable but
not catastrophic amount.

@@

The challenge was heavily simplified, with the intention of testing whether
effective bin-assignments with limited photometry is possible even in theory,
and of generating as many potential methods for the task as possible.  With
both these achieved, future directions for simulating the problem are clear.
The most obvious is limited spectroscopic completeness and size.  Realistic
training data sets will be far smaller than the millions of objects we used here,
and the difficulty of measuring spectra for faint galaxies will make the them
incomplete and highly unrepresentative. Machine learning methods will be particularly
affected by the latter problem.  Future challenges must explore these important limitations
using simulated incompleteness, and use realistic photo-z methods to compute metrics.

@@

Method implementations used in the challenge may be found in the \textsc{grand\_merge} branch
of the challenge repository \url{http://github.com/LSSTDESC/tomo_challenge}.

@@

We combine the convolutional layers with dimensionality reduction (pooling layers) and different
activation functions. Here, we used Autokeras to optimize a general convolutional architecture,
using the galaxies' magnitudes and colours to assign them to redshift bins, which were chosen such
that an equal number of galaxies were assigned to each.

@@

In addition to this base functionality, the SimpleSOM method includes a
mechanism for distilling $N_{\rm cell}$ SOM cells into $N_{\rm group}<N_{\rm
cell}$ groups of cells, while maintaining optimal redshift resolution. This is
done by invoking a hierarchical clustering of SOM cells \citep[see Appendix B of
][]{Wright/etal:2020a}, where in this case we combine cells based on the
similarity of the cells' $n(z)$ moments (mean, median, and normalised median absolute
deviation from median) using complete-linkage clustering. The result of this
procedure is the construction of fewer discrete groups of training and reference
data, while not introducing pathological redshift-distribution broadening (as
happens when, e.g., training using a lower resolution SOM and/or clustering
cells based on distances in colour-magnitude space).

@@

We chose the learning algorithm for both methods to minimize the reciprocal of
the figure of merit $F$, where $F$ is defined in \ref{eq:fom}. We implement
these models using the JAX-based Flax neural network library \cite{jax}.
